---
title: "Untitled"
author: "Hoang-Chau Luong"
date: '2022-09-12'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

# **Easy**

## **1**

State the three motivating criteria that define information entropy. Try to express each in your own words.

**Answer**

Insight of information entropy is "How much is our uncertainty reduced by learning an outcome? 3 criteria:

1. The measure of uncertainty should be continuous
2. The measure of uncertainty should increase as the number of possible events increases -> Sample set larger -> uncertainty increases
3. The measure of uncertainty should be additive

## **2**

:::quesion
>Suppose a coin is weighted such that, when it is tossed and lands on a table, it comes up heads 70% of the time. What is the entropy of this coin?
:::

```{r}
p_logp <- function(p){
  if(p == 0) return(0)
  p * log(p)
}

calc_entropy <- function(x){
  avg_logprob <- sum(map_dbl(x, p_logp))
  -1 * avg_logprob
}

probs <- c(0.7, 0.3)

calc_entropy(probs)
```

## **3**

:::question
> Suppose a four-sided die is loaded such that, when tossed onto a table, it shows “1” 20%, “2” 25%, “3” 25%, and “4” 30% of the time. What is the entropy of this die?
:::

```{r}
probs <- c(0.2, 0.25, 0.25, 0.3)
calc_entropy(probs)
```

## **4**

:::question
>Suppose another four-sided die is loaded such that it never shows “4.” The other three sides show equally often. What is the entropy of this die
:::

```{r}
probs <- c(1/3, 1/3, 1/3, 0)
calc_entropy(probs)
```

# **Medium**

## **1**

:::question
>Write down and compare the definitions of AIC and WAIC. Which of these criteria is most general? Which assumptions are required to transform the more general criterion into a less general one?
:::

$$
\text{lppd} = \sum_{i}^{n}\text{log}
\int\text{p}(\text{y}_{i}|\theta)\text{p}_{post}(\theta)d\theta
$$

```{r}
elpd <- function(object, method="LOO"){
  
  object$output[-1] <- paste(object$output$outFilePath,object$output[-1],sep="")
  
  if(is.null(object$output$CPO))
    stop("Please specify the argument 'output_CPO = TRUE' in BayesSUR()!")
    
  if(toupper(method) == "LOO"){
    elpd <- sum(log(read.table(object$output$CPO)))
    names(elpd) <- "elpd.loo"
  }else if(toupper(method) == "WAIC"){
    elpd <- sum(read.table(object$output$WAIC))
    names(elpd) <- "elpd.waic"
  }else{
    stop("Please give the correct method name!")
  }
  
  return(elpd)
}
```






















